Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape

### 2024-04-24 paper/paper.qmd
final paper updates

To further explore the contextual relationship between each word, @fig-word-networks creates word networks of all nine SOC texts. These networks capture how often each word frequents in these SOC texts and how closely each correlates with one another, whether as single words (unigrams) or pairs of words (bigrams). When combined all SOC texts, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times ([Appendix]'s @tbl-count). This further correlates to the non-linear nature of SOC literature and their exploration of cognitive incoherence. In terms of correlation, ignoring personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include those like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others ([Appendix]'s @tbl-correlation). Observable from  @fig-word-networks, with a correlation and a word count of at least 0.8 and 400, \textit{people, moment, life} and \textit{mind} are the most predominant word combinations throughout all 9 SOC texts.

@@

The AFINN index in @fig-sentiment-number portrays a similar trend with its more quantitative sentiment analysis. On a scale of -5 to 5, SOC literature from these five authors mostly carry a negative sentiment, observable in the significant difference in height between the negative trendlines and the positive trendlines. Eliot's works are the most negative with 37% of words being -2 on the emotional scale between -5 and 5. Respectively, Woolf, Proust and Joyce tail beind at 33%, 32% and 31%, averaging out to be -2.2 on the emotional scale between -5 and 5. While a large percentage Mansfield's works are positive, her highest percentage is negative on the AFINN index, at 24% or a -2.2 negative score between -5 and 5.

@@

While sharing the same existential quest in their SOC literature, through this gendered analysis and @fig-sentiment-category, we can see discrepancies between male and female authors, especially in terms of shared linguistic patterns and mental health themes. @gender poses that in terms of genders, writers-or here SOC writers-tend to write about affects or mental health themes differently. While females tend to have reservations over their emotions or are similarly positive, males tend to be more open and individualistic due to less social constraints at the time, which will be further discussed in [Discussion] below.

@@

As scholars Bernini and Fernyhough mentioned, the "mysterious nature of the conscious" is "a simultaneously fragmented and unified mental realm" [@resampling], this analysis has also reflected that sentiment in SOC literature. Most of the texts express the desire to replicate the inner workings of the human mind textually, attempting to capture the complexity of thought processes and emotions. With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in an authentic manner, deeply portraying then-taboo topics such as existential angst, anxiety and cognitive dissonance.

>>>>

To further explore the contextual relationship between each word, @fig-word-networks creates word networks of all nine SOC texts. These networks capture how often each word frequents in these SOC texts and how closely each correlates with one another, whether as single words (unigrams) or pairs of words (bigrams). When combined all SOC texts, temporal nouns like \textit{time, day, life, moment, night} appear the most, at, respectively, 870, 597, 567, 502 and 376 times ([Appendix]'s @tbl-count). This further correlates to the non-linear nature of SOC literature and their exploration of cognitive incoherence. In terms of correlation, ignoring personal names like \textit{swann, odette or mme}, bigrams with a correlation of 1 include word pairs like \textit{dark, love}; \textit{day, heart}; \textit{moment, life}; \textit{mind, life}; \textit{moment, mind}, among others ([Appendix]'s @tbl-correlation). Observable from  @fig-word-networks, with a correlation and a word count of at least 0.8 and 400, \textit{people, moment, life} and \textit{mind} are the most predominant word combinations throughout all 9 SOC texts.

@@

The AFINN index in @fig-sentiment-number portrays a similar trend. On a scale of -5 to 5, SOC literature from these five authors mostly carry a negative sentiment, observable in the significant difference in height between the negative trendlines and the positive trendlines. Eliot's works are the most negative with 37% of words being -2 on the emotional scale between -5 and 5. Respectively, Woolf's, Proust's and Joyce's tail beind at 33%, 32% and 31%, averaging out to be -2.2 on the emotional scale between -5 and 5. While a large percentage Mansfield's works are positive, 24% of her words-the largest percentage of hers-are negative, earning a -2.2 negative score between -5 and 5 in the AFINN index.

@@

Through this gendered analysis and @fig-sentiment-category, we can see discrepancies between male and female SOC authors, especially in terms of shared linguistic patterns and mental health themes. @gender poses that in terms of genders, writers-or here SOC writers-tend to write about affects or mental health themes differently. While females tend to have reservations over their emotions or are similarly positive, males tend to be more open and individualistic due to less social constraints at the time, which will be further discussed in [Discussion] below.

@@

As scholars Bernini and Fernyhough mentioned, the "mysterious nature of the conscious" is "a simultaneously fragmented and unified mental realm" [@resampling], this analysis has also reflected that sentiment in SOC literature. Most of the texts express the desire to textually replicate the inner workings of the human mind, attempting to capture the complexity of thought processes and emotions. With these inner-workings revealed to be mostly negative and existential, SOC authors have explored their identity and worldview in an authentic manner, deeply portraying then-taboo topics such as existential angst, anxiety and cognitive dissonance.
